### 2016-04-22 paper.tex
Add base template for paper.

\documentclass{article}

\title{APPM 3310 Project\\\Large{Static Analysis of Truss Systems}}
\author{Sammy Chang\and{}Ryan Marizza\and{}Michael Shannon}
\date{Friday, April 29, 2016}

\usepackage{hyperref}
\hypersetup{colorlinks=true, urlcolor=blue, linkcolor=blue, citecolor=blue}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{graphicx}

\usepackage{abstract}
\renewcommand{\abstractname}{}
\renewcommand{\absnamepos}{empty}
\renewcommand{\abstracttextfont}{\normalfont\small\bfseries}

\usepackage[citestyle=authoryear]{biblatex}
\addbibresource{paper.bib}

\begin{document}

\maketitle

\begin{abstract}
    Abstract goes here.
\end{abstract}

\section{Introduction}

\section{Mathematical Formulation}

\section{Examples and Numerical Results}

\section{Discussion}

\section{Conclusion}

\nocite{*}
\printbibliography{}

\end{document}