### 2016-04-25 paper.tex
Add Introduction from Sammy Chang.

\documentclass{article}

\title{APPM 3310 Project\\\Large{Static Analysis of Truss Systems}}
\author{Sammy Chang\and{}Ryan Marizza\and{}Michael Shannon}
\date{Friday, April 29, 2016}

\usepackage{hyperref}
\hypersetup{colorlinks=true, urlcolor=blue, linkcolor=blue, citecolor=blue}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{listings}

\usepackage{abstract}
\renewcommand{\abstractname}{}
\renewcommand{\absnamepos}{empty}
\renewcommand{\abstracttextfont}{\normalfont\small\bfseries}

\usepackage[citestyle=authoryear]{biblatex}
\addbibresource{paper.bib}

\begin{document}

\maketitle

\begin{abstract}
This report covers application of Linear Algebra in the Finite Element Method.
More specifically we delve into the Direct Stiffness Method (DSM) which is one
common method to apply FEM. Our code is able to solve 2D and 3D truss systems
using DSM. DSM is essentially based on three knowns: forces, displacements and
elastic modulus'. The scope of the project only covers 2D trusses, but the
code does have the ability to solve 3D systems.
\end{abstract}

\section{Introduction}
The truss analyzer implements the DSM by simplifying and idealizing a given
truss system. By that, all of the joints are converted to frictionless pin
joins, and we will ignore the bending of individual truss members. Our
idealization for most cases is satisfactory for an elementary truss analysis.
Therefore we will only be observing the axial forces and the displacements at
the pin joints. Our truss analyzer "converts" the pins and members to a node
and an element, respectively under the Cartesian coordinate system.
Furthermore, we are able to apply external forces only at the joints. This
will allow us to simulate displacement in all the nodes of the truss system.
The analyzer works by inputting a truss system, the code then identifies
individual elements. When the elements are identified, the truss system is
disassembled at the nodes. The elements are further analyzed using stiffness
equations dependent on the geometry and the elastic moduli of the elements.
Once those elements have been analyzed, the code re-assembles the structure
with the stiffness of the elements. The element stiffness matrix is then
modified by applying the displacements and externally applied forces at the
nodes. A final matrix is obtained (global stiffness matrix) which can be
solved using the displacement and force vectors by applying the Cholesky
decomposition. Using DSM allows us to analyze complex truss structures as well
as statically indeterminate structures.

\section{Mathematical Formulation}

Most of the following formulation follows directly from
\parencite{felippa2004ch2} and \parencite{felippa2004ch3}.

In this paper the accompanying tools use the direct stiffness method.  This
method is based on using known forces and displacements as well as the modulus
of elasticity of the structure members to back out displacements.  In this
way, it is unlike the direct force method which can only solve statically
determinate systems as the direct stiffness method can solve indeterminate
systems as well by properly splitting the stress among the truss members.

The direct stiffness method involves 8 steps.  The first two are purely
conceptual and thus are never really performed and the last one is a post
processing step performed after a solution is found.  These eight steps are
outlined below (taken from Figure 2.5 of \cite{felippa2004ch2}).
\begin{enumerate}[noitemsep]
    \item Disconnection
    \item Localization
    \item Element Formation
    \item Globalization
    \item Merge
    \item Application of Boundary Conditions
    \item Solution
    \item Recovery of Derived Quantities
\end{enumerate}

The disconnection step is essentially just looking at each truss member
(element) individually while localization is just looking at member based
coordinate systems.  Therefore, this paper will skip directly to the first
step requiring actual work, \emph{element formation}.

\subsection{Element Formation}

The eventual goal is to find the master stiffness equation
\begin{equation}
    \mathbf{f}=K\mathbf{u} \label{eq:master_stiffness_equation}
\end{equation}
where $\mathbf{f}$ is the vector of external forces, $\mathbf{u}$ the vector
of displacements, and $K$ the linear relation between the two.  This equation
is in a global coordinate frame.  $K$ is $n\times{}n$ where $n$ is the number
of nodes (endpoint of one or more members) times the number of dimensions
which for the remainder of this document will be taken as 2.

Before the master stiffness equation can be written a relation between
displacement and external force for each element must be developed.  To do
this, the elements of the truss are treated as springs and therefore governed
by Hook's law.
\begin{equation}
    f=k_su \label{eq:hooks_law}
\end{equation}
where $f$ is the force, $u$ is the displacement, and $k_s$ is the stiffness of
the spring, which for an element of constant cross sectional area $A$, of
length $L$ and made of a material with elastic modulus $E$ is given by:
\begin{equation}
    k_s=\frac{EA}{L} \label{eq:stiffness_coefficient}
\end{equation}

By making the simplification that nodes cannot impart any torque on elements
the only equation governing an element is Hook's law.  In the element centered
coordinate frame the one dimensional Hook's law, Eq.~(\ref{eq:hooks_law}) is
enough.  However, in order to convert this into the global coordinate frame a
full dimensional equation is needed.  This equation will be designated as the
local stiffness equation for element $e$:
\begin{equation}
    \overline{\mathbf{f}}^e=\overline{K}^e\overline{\mathbf{u}}^e
    \label{eq:local_stiffness_equation}
\end{equation}
This equation is in a member centric coordinate frame with the origin placed
on the first node of the element and the $x$-axis along the element, pointing
to the 2nd node.  The direction of all other axes are unimportant.  Therefore,
in the 2-dimensional case $\overline{K}^e$ becomes:
\begin{equation}
    \overline{K}^e=\frac{EA}{L}\begin{bmatrix*}[r]
        1&0&-1&0\\
        0&0&0&0\\
        -1&0&1&0\\
        0&0&0&0
    \end{bmatrix*} \label{eq:kbar}
\end{equation}
The 3-dimensional case can be easily obtained by adding 2 zero rows and 2 zero
columns.

\subsection{Globalization}

Before the master stiffness matrix can be formed,
Eq.~(\ref{eq:local_stiffness_equation}) must be converted into the global
coordinate frame equation:
\begin{equation}
    \mathbf{f}^e=K^e\mathbf{u}^e\label{eq:global_stiffness_equation}
\end{equation}
which relates displacement to force for a single element $e$ in the global
coordinate frame.  Let $\mathbf{x}$ be a vector in the global coordinate
system and $\overline{\mathbf{x}}$ be the equivalent vector in the element
local coordinate frame.  Then the relation between the two is given by
\parencite{olver2006}:
\begin{equation}
    S\overline{\mathbf{x}}=\mathbf{x}\qquad\text{where}\qquad
    S=\left\{\mathbf{v}_1,\mathbf{v}_2,\ldots,\mathbf{v}_n\right\}
\end{equation}
With the $\mathbf{v}$ vectors forming a basis for the element centered
coordinate system.  In particular, $v_1$ is the unit vector along the element.
If the other columns of $S$ are chosen such that $S$ is orthonormal then
$S^{-1}=S^{T}$.  Therefore, the reverse transformation is:
\begin{equation}
    \overline{\mathbf{x}}=S^{-1}\mathbf{x}=S^T\mathbf{x}
\end{equation}
Noting that $\mathbf{u}$ and $\mathbf{f}$ are stacked vectors that rotate
though each coordinate twice the appropriate transformation matrix for element
$e$ becomes:
\begin{equation}
    T^e=\begin{pmatrix}
        S^T&O\\O&S^T
    \end{pmatrix}
\end{equation}
Therefore, Eq.~(\ref{eq:global_to_local}) shows the relation between the
global coordinate frame and the local coordinate frame for the force and
displacement vectors.
\begin{equation}
    \overline{\mathbf{u}}^e=T^e\mathbf{u},\qquad
    \overline{\mathbf{f}}^e=T^e\mathbf{f}
    \label{eq:global_to_local}
\end{equation}
By combining Eq.~(\ref{eq:local_stiffness_equation},
\ref{eq:global_stiffness_equation}, \ref{eq:global_to_local}) the global
stiffness matrix (for element $e$), $K^e$ can be related to the local
stiffness matrix $\overline{K}^e$ by:
\begin{equation}
    K^e=(T^e)^T\overline{K}^eT^e
\end{equation}

\subsection{Merge}

The master stiffness equation Eq.~(\ref{eq:master_stiffness_equation}) relates
the displacement of every node, stored in $\mathbf{u}$ to every external force
in vector $\mathbf{f}$ using the linear transformation $K$.  The assembly of
each $K^e$ into $K$ is governed by \parencite{felippa2004ch3}:
\begin{enumerate}[noitemsep]
    \item Compatibility of displacements at any given node.
    \item Equilibrium of internal and external forces.
\end{enumerate}
The typical method of handling assembly of the global stiffness equations,
Eq.~(\ref{eq:global_stiffness_equation}), is to sum the forces and sum the
$K^e$ matrices by first expanding them to the size of $K$ by adding rows and
columns of zeros in the appropriate locations.  In the case of the computer
code that accompanies this document \parencite{shannon2016} the method in
section 3.5.1 of \cite{felippa2004ch3} is used.  This is know as
\emph{Assembly by Freedom Pointers}.  The algorithm is governed by
Eq.~(\ref{eq:global_to_master}) copied from \cite{felippa2004ch3} and fills $K$
by iterating through each element $e$ and then through the values in $K^e$
where $d=4$ in the 2-dimensional case and $d=6$ in the 3-dimensional case.
Let the number of elements in the truss be given by $m$.

\begin{minipage}{\textwidth}
\begin{equation*}
    K_{pq}=\sum_{e=1}^{m}K_{ij}^e
\end{equation*}
\begin{equation}
    \text{for}\label{eq:global_to_master}
\end{equation}
\begin{equation*}
    i=1,2,\ldots,d,\quad{}j=1,2,\ldots,d\qquad{}
    p=EFT^e(i),\quad{}q=EFT^e(j)
\end{equation*}
\vspace{0.1em}
\end{minipage}

The \emph{element freedom table} ($EFT^e$) is unique for each element $e$ and
gives the row/column indices where the values in $K^e$ are to be added into
$K$.  The specific algorithm to get $EFT^e$ is better explained in code and
can be found at \lstinline$libs/element_freedom_table.m$ in
\cite{shannon2016}.  In reality, $K$ is initialized to the zero matrix.  The
force vector $\mathbf{f}$ is simply constructed by summing the known (applied)
external forces in the proper indices which can also be obtained from the
\emph{element freedom table}.

\subsection{Application of Boundary Conditions}

Eq.~(\ref{eq:master_stiffness_equation}) is singular because while the truss
is \emph{hopefully} self constrained at this point it is not constrained in
space.  This is because Eq.~(\ref{eq:master_stiffness_equation}) does not yet
take into acount the fact that $\mathbf{u}$ is under some restrictions due to
constraints on the physical truss.  For the purpose of this paper, and the
accompanying software \parencite{shannon2016}, constraints will be limited to
one of the principal ($xyz$) dimensions.  Therefore, some of the entries of
$\mathbf{u}$ are forced to be zero and in these locations the force is
unknown.  The normal way to handle this is to simply remove the associated
rows and columns from $\mathbf{f}$, $\mathbf{u}$, and $K$.  However, in a
software package this either requires special logic when implementing
Eq.~(\ref{eq:global_to_master}) or requires reallocating a new $K$ matrix
after the fact.

Therefore, in this paper the computer friendly method outlined in section
3.5.2 of \cite{felippa2004ch3} is used.  This method involves setting the
diagonals of $K$ that correspond to the constrained parts of $\mathbf{u}$ to 1
and the corresponding values in $\mathbf{f}$ to 0.  This forces those elements
of $\mathbf{u}$ to solve to 0 displacement.  This is a simple $O(n)$ operation
instead of the higher order delete of rows and columns.

\subsection{Solution}

Finally, Eq.~(\ref{eq:master_stiffness_equation}) is non-singular and thus can
be solved by any number of methods.  The result is the column vector of
displacements $\mathbf{u}$.  This gives the displacements of each node and
therefore the displacement of the endpoints of each element.  This is enough
to plot the new layout of the truss under load.

\subsection{Recovery of Derived Quantities}

\section{Examples and Numerical Results}

\section{Discussion}

\section{Conclusion}

% \nocite{*}
\printbibliography{}

\end{document}